---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load the activity data into dataframe "activity".
```{r , echo=TRUE}
setwd("~/Src/R/Repro_Research_coursera/RepData_PeerAssessment1/")
activity <- read.csv("activity.csv", header = TRUE)
summary(activity)
```
The summary of the data reveals that there are many NA's.  remove incomplete (NA) rows from subsequent analysis.

```{r, echo=TRUE}
good <- complete.cases(activity)
summary(activity[good, ])
```


## What is mean total number of steps taken per day?
First find the total steps in each day
```{r, echo=TRUE}
mean(activity[good, "steps"])
day_activity <- tapply(activity[good, "steps"], activity[good, "date"], sum )

day_activity
```
Next find the mean of the days with good values

```{r , echo=TRUE}
mean(day_activity[complete.cases(day_activity)])
median(day_activity[complete.cases(day_activity)])
hist(day_activity[complete.cases(day_activity)], br=10, xlab="daily steps",ylab="days", main="Histogram of total steps taken each day")
```


## What is the average daily activity pattern?

```{r , echo=TRUE}
time_activity <- tapply(activity[good, "steps"], activity[good, "interval"], sum )
plot(time_activity, xlab="5-min intervals",ylab="steps taken", main="Time series of steps taken averaged across days")
```

What 5 minute interval gave the most steps?
```{r , echo=TRUE}
which.max(time_activity)
```
The most activity in a 5 minute interval occurs at 8:35 in the morning.

## Imputing missing values
replace NA values with the mean number of steps for that time interval.
```{r , echo=TRUE}
activityTmp <- activity
dim(time_activity)
str(levels(activity$date))
num_dates <- 61
new_steps <- rep(time_activity, times = num_dates)
activityTmp$steps <- new_steps
```
Now replace the NA's with the values in new_steps
```{r , echo=TRUE}
activityNew <- activity
activityNew[!good, ] = activityTmp[!good, ]
summary(activityNew)
```



## Are there differences in activity patterns between weekdays and weekends?
